Contrastive Optimization -- Preliminaries

### 2024-10-23 paper/main.tex
ce/ace now have representative comparison

\begin{theorem}\label{ablatedce}
	We can integrate background invariance to the definition of cross-entropy using the following formulation:
	\begin{gather}
		\mathcal{J}(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} + \sum |(1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}|\right)} \right)
	\end{gather}
\end{theorem}
\begin{proof} Using the definition of HiResCAMs, ContrastiveCAMs and the elementwise difference setup from \ref{classprobdiff}, we have:
	\begin{align}
		d &= [s_i - s_c]_i \\
		&= \left[\sum^{D_1,D_2}_{i,j,k} \tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - \sum^{D_1,D_2}_{c,j,k} \tilde{\mathcal{\bm{A}}}_{c,j,k}^{\text{HiResCAM}}\right] \\
		&= \left[ -\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} \right]_i
	\end{align}
	As an aside: we note that when $i = c$, the contrastive map is the zero matrix $\in \mathbb{R}^{D_1 \times D_2}$, and instead of $|C| - 1$ contrastive maps we now obtain $|C|$ contrastive maps. Using our result from \ref{classprobdiff}, we have:
	\begin{gather}
		\sigma_c(\bm{s}) = \frac{1}{\sum_i e^{d_i}} = \frac{1}{\sum_i \text{exp}({-\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}})}
	\end{gather}
	For a target class $c$, cross-entropy loss is defined as:
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \sigma_c(\bm{s}) \right) = -\log \left( \frac{1}{\sum_i \text{exp}({-\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}})} \right)
	\end{gather}
	We can break down each contrastive map to foreground and background components:
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} - \sum (1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}\right)} \right)
	\end{gather}
	The current loss formulation motivates improving contrast using either the foreground or the background. We penalize contrast on the background, motivating learning invariance to noise in the form of the background.
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} + \sum |(1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}|\right)} \right)
	\end{gather}
	This gives us our ablated cross-entropy formulation.
\end{proof}

>>>>

\begin{theorem}\label{ablatedce}
	We can integrate background invariance to the definition of cross-entropy using the following formulation:
	\begin{gather}
		\mathcal{J}(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} + \sum |(1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}|\right)} \right)
	\end{gather}
\end{theorem}
\begin{proof} Using the definition of HiResCAMs, ContrastiveCAMs and the elementwise difference setup from \ref{classprobdiff}, we have:
	\begin{align}
		d &= [s_i - s_c]_i \\
		&= \left[\sum^{D_1,D_2}_{j,k} \tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{HiResCAM}} - \sum^{D_1,D_2}_{j,k} \tilde{\mathcal{\bm{A}}}_{c,j,k}^{\text{HiResCAM}}\right]_i \\
		&= \left[ -\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} \right]_i
	\end{align}
	As an aside: we note that when $i = c$, the contrastive map is $\bm{0} \in \mathbb{R}^{D_1 \times D_2}$, and instead of $|C| - 1$ contrastive maps we now obtain $|C|$ contrastive maps. \newline \\
	Using our result from \ref{classprobdiff}, we have:
	\begin{gather}
		\sigma_c(\bm{s}) = \frac{1}{\sum_i e^{d_i}} = \frac{1}{\sum_i \text{exp}({-\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}})}
	\end{gather}
	For a target class $c$, cross-entropy loss is defined as:
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \sigma_c(\bm{s}) \right) = -\log \left( \frac{1}{\sum_i \text{exp}({-\sum^{}\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}})} \right)
	\end{gather}
	We can break down each contrastive map to foreground and background components:
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} - \sum (1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}\right)} \right)
	\end{gather}
	The current loss formulation motivates improving contrast using either the foreground or the background. We penalize contrast on the background, motivating learning invariance to noise in the form of the background.
	\begin{gather}
		H(\theta, \mathcal{D}) = -\log \left( \frac{1}{\sum_i \text{exp}\left({-\sum^{}h \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)} + \sum |(1-h) \odot \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, i)}}|\right)} \right)
	\end{gather}
	This gives us our ablated cross-entropy formulation.
\end{proof}
